### 2012-02-10 trunk/GUIErrorDetector/paper/introduction.tex
address the first round of comments

The single-GUI-thread rule can be easily violated since a spawned non-UI thread often needs to update
the GUI  after its  task is finished.
In practice, invalid thread access errors are \textit{frequent}, \textit{severe}, and \textit{hard} to debug.
%and \textit{difficult to avoid}.

Take the popular Standard Widget Toolkit (SWT) GUI framework as an example. 
The invalid thread access error is one of the top 3 bugs in developing a SWT application, and is
the source of many relevant concurrency bugs~\cite{top3bugs}.
A Google search for ``SWTException:Invalid thread access''  returns over 11,800 entries,
consisting of numerous bug reports, forum posts, and mailing list threads on
this problem. Eclipse, the IDE for Java development, is built on top of SWT\@.
Searching for ``SWTException:Invalid thread access'' in Eclipse's bug repository
and discussion forum returns over 2732 bug reports and 351 distinct threads, respectively. 
We manually studied all 156 distinct \textit{confirmed} bug reports, and 
found this error has been reported in at least 20 distinct Eclipse projects
and 40 distinct Eclipse components. Even after over 10 years of active development,
a recent release of Eclipse still contains this error (bug id: 333533, reported in 01/2011).
 In addition, the invalid thread access error
is severe. It is user-perceivable; it cannot be recovered by the program itself,
and it often terminates the whole application. In many circumstances as described in the bug reports,
users can only restart the application to recover from the error.
Furthermore, many reported  errors are non-trivial to diagnose and fix.
Developers usually need non-local reasoning to find the specific
UI interactions that can trigger the bug;  it took developers 2 years
to fix bug 51757 and verify the patch.

@@

To ensure that GUIs behave correctly, developers must prevent or detect invalid thread
access errors. Currently, no effective techniques exist. %must be developed.% to alleviate this problem.

%Software testing, the most widely used method to ensure program correctness,
%is insufficient to detect such errors. 
It is infeasible for testing to cover the enormous space of possible interactions
with a GUI. Each sequence of GUI events can result in
a different state, and each GUI event may need to be evaluated in all of
these states. A well-developed software system like Eclipse
often has a well-designed test suite that achieves fairly high statement coverage;
but many paths executed by bug-triggering UI sequences are still not covered.

Stylized coding patterns are also inadequate. One possible rule is to always
access GUI objects via asynchronized message passing, to ensure a GUI object is accessed in
the UI thread. For example, a developer could have prevented the bug in Figure~\ref{fig:androiderror}
by wrapping line 222 inside a \CodeIn{post} message passing method\footnote{The
\CodeIn{post} method in class \CodeIn{android.widget.View} is a
a standard way to send asynchronized messages to the UI-thread.}, as follows:

@@

Such an approach is desirable for accesses from non-UI threads, but it is unnecessary
and dangerous for accesses from the UI thread. Asynchronized message passing offers
no timing guarantee, so a GUI object may have already been
disposed before the message sent to it arrives, causing other bugs. 
In our evaluation on real-world programs, we found a simple analysis that
requires GUI operations to be in a message issues
an unacceptable number of warnings. %false positives and redundant warnings.

@@

This paper uses static analysis to find potential GUI errors.
Static analysis has two advantages compared to dynamic approaches such as
 testing. First, a static analysis can explore paths of the program without
executing the code, and without the need for a test suite.
Second, a static analysis can verify the code: if it
reports no warnings, the code is guaranteed to be bug-free.

\begin{figure}[t]
\begin{CodeOut}
\begin{alltt}
   org.mozilla.gecko.gfx.LayerView.<init>(Context;LayerController)
-> android.opengl.GLSurfaceView.setRenderer(GLSurfaceView\$Renderer;)
-> java.lang.Thread.start()
-> android.opengl.GLSurfaceView\$GLThread.run()
-> android.opengl.GLSurfaceView\$GLThread.guardedRun()
-> org.mozilla.gecko.gfx.LayerRenderer.onSurfaceChanged(GL10;II)
-> org.mozilla.gecko.gfx.LayerView.setViewportSize(IntSize;)
   ... 
-> android.view.ViewRoot.recomputeViewAttributes(View;)
-> android.view.ViewRoot.checkThread()
\end{alltt}
\end{CodeOut}
%\hspace{5mm}\CodeIn{\scriptsize{org.mozilla.gecko.gfx.LayerView.<init>(Context;LayerController)}}\\
%$\rightarrow$\CodeIn{ \scriptsize{android.opengl.GLSurfaceView.setRenderer(GLSurfaceView\$Renderer;)}}\\
%$\rightarrow$\CodeIn{ \scriptsize{java.lang.Thread.start()}}\\
%$\rightarrow$\CodeIn{ \scriptsize{android.opengl.GLSurfaceView\$GLThread.run()}}\\
%$\rightarrow$\CodeIn{ \scriptsize{android.opengl.GLSurfaceView\$GLThread.guardedRun()}}\\
%$\rightarrow$\CodeIn{ \scriptsize{org.mozilla.gecko.gfx.LayerRenderer.onSurfaceChanged(GL10;II)}}\\
%$\rightarrow$\CodeIn{ \scriptsize{org.mozilla.gecko.gfx.LayerView.setViewportSize(IntSize;)}}\\
%\hspace{5mm}\CodeIn{...}\\
%$\rightarrow$\CodeIn{ \scriptsize{android.view.ViewRoot.checkThread()}}
\tinystep
\vspace*{-3.0ex} \Caption{{\label{fig:report} Our tool reports
a method call chain that reveals the potential error in
Figure~\ref{fig:androiderror}. $\rightarrow$ represents
the call relationship
between methods, and \CodeIn{checkThread} is an Android library
method  that checks whether the current thread is the
event dispatching thread before accessing a GUI object.
8 more methods in the call chain, shown as
\CodeIn{\small{...}} above, are omitted for brevity.
}} %\vspace{-5mm}
\end{figure}

@@

Our static analysis is independent of the call graph construction algorithm.
%Given a sound call graph, our analysis is sound in that it does
%not miss true positives.
However, modern GUI applications tend to use reflection, and
in the presence of reflection, existing call graph construction algorithms such as RTA~\cite{rta}
and k-CFA~\cite{kcfa} fail to build a good call graph.
To alleviate this problem, we use an algorithm to build reflection-aware call graph, and
also compare its usefulness with existing call graph construction algorithms in
our experiments. Static analysis may report false positives due to its
conservative nature, or multiple warnings that actually correspond
to the same error. To address such limitations, we devise a set
 of error filters to remove likely false positives and redundant warnings.

@@

We implemented an invalid-thread-error detection tool that supports
four popular GUI frameworks: SWT, Eclipse plugin environment,
Swing, and Android. Swing and SWT are the
two dominant GUI frameworks for desktop Java applications.
Eclipse is the de-facto standard IDE for Java. Android is the \#1 platform for
mobile applications with market share 56\% as of Sep 2011.
Although our technique is applicable to any GUI
framework with the single-GUI-thread rule, each framework
has its own definition of \textit{UI thread} and \textit{program 
entry points}. Thus, our implementation is parameterized with respect to
those framework-specific parts.

Our tool works in a fully automatic, push-button manner,
and scales to realistic programs.
We evaluated our implementation on \subnum programs comprising \totaloc LOC.
The experimental results demonstrate
that: our technique is effective (it found \bugs real-world errors
from four supported frameworks and produced only \falses false positive warnings);
our reflection-aware call graph construction algorithm helps in
finding errors in Android applications;
and our proposed filters significantly reduce the number of warnings.
%false positives and redundant warnings.

@@

\item \textbf{Technique.} We formulate finding
the invalid thread access error as a call graph reachability problem,
and present a general error detection technique.
In addition, we use a reflection-aware
call graph construction algorithm. (Section~\ref{sec:technique}).

>>>>

The single-GUI-thread rule can be easily violated; since a spawned non-UI thread often needs to update
the GUI  after its  task is finished.
In practice, invalid thread access errors are \textit{frequent}, \textit{severe}, and \textit{hard to debug}.
%and \textit{difficult to avoid}.

Take the popular Standard Widget Toolkit (SWT) GUI framework as an example. 
Invalid thread access is one of the top 3 bugs in developing a SWT application, and is
the source of many concurrency bugs~\cite{top3bugs}.
A Google search for ``SWTException:Invalid thread access''  returns over 11,800 entries,
consisting of numerous bug reports, forum posts, and mailing list threads on
this problem. Eclipse, the IDE for Java development, is built on top of SWT\@.
Searching for ``SWTException:Invalid thread access'' in Eclipse's bug repository
and discussion forum returns over 2732 bug reports and 351 distinct threads, respectively. 
We manually studied all 156 distinct \textit{confirmed} bug reports, and 
found this error has been confirmed in at least 20 distinct Eclipse projects
and 40 distinct Eclipse components. Even after over 10 years of active development,
a recent release of Eclipse still contains this error (bug id: 333533, reported in January 2011).
 In addition, the invalid thread access error
is severe. It is user-perceivable, it cannot be recovered by the program itself,
and it often terminates the whole application. In many circumstances as described in the bug reports,
users must restart the application to recover from the error.
Furthermore, many reported  errors are non-trivial to diagnose and fix.
Developers usually need non-local reasoning to find the specific
UI interactions that can trigger the bug;  it took developers 2 years
to fix Eclipse bug 51757 and verify the patch.

@@

To ensure that GUIs behave correctly, developers must prevent or detect invalid thread
access errors. Current techniques are not effective. %must be developed.% to alleviate this problem.

%Software testing, the most widely used method to ensure program correctness,
%is insufficient to detect such errors. 
It is infeasible for testing to cover the enormous space of possible interactions
with a GUI. Each sequence of GUI events can result in
a different state, and each GUI event may need to be evaluated in all of
these states. A  software system like Eclipse
often has a test suite that achieves fairly high statement coverage,
but many paths executed by bug-triggering UI sequences are still not covered.

Stylized coding patterns are also inadequate. One possible rule is to always
access GUI objects via asynchronous message passing, to ensure a GUI object is accessed in
the UI thread. For example, a developer could have prevented the bug in Figure~\ref{fig:androiderror}
by wrapping line 222 inside a \CodeIn{post} message-passing method\footnote{The
\CodeIn{post} method in class \CodeIn{android.widget.View} is a
a standard way to send asynchronized messages to the UI-thread.}, as follows:

@@

Such an approach is desirable for accesses from non-UI threads, but it is unnecessary.
In our evaluation on real-world programs, we found a simple analysis that
requires GUI operations to be in a message issues
an unacceptable number of warnings. %false positives and redundant warnings.
Furthermore, this approach is dangerous for accesses from the UI thread. Asynchronized message passing offers
no timing guarantee, so a GUI object may have already been
disposed before the message sent to it arrives, causing other bugs.

@@

\smallskip

\textbf{Our approach: static analysis.}
This paper uses static analysis to find potential GUI errors.
Static analysis has two advantages compared to dynamic approaches such as
 testing. First, a static analysis can explore paths of the program without
executing the code, and without the need for a test suite.
Second, a static analysis can verify the code: if a sound static analysis
reports no warnings, the code is guaranteed to be bug-free.

\begin{figure}[t]
\begin{CodeOut}
\begin{alltt}
   org.mozilla.gecko.gfx.LayerView.<init>(Context;LayerController)
-> android.opengl.GLSurfaceView.setRenderer(GLSurfaceView\$Renderer;)
-> java.lang.Thread.start()
-> android.opengl.GLSurfaceView\$GLThread.run()
-> android.opengl.GLSurfaceView\$GLThread.guardedRun()
-> org.mozilla.gecko.gfx.LayerRenderer.onSurfaceChanged(GL10;II)
-> org.mozilla.gecko.gfx.LayerView.setViewportSize(IntSize;)
   ... 
-> android.view.ViewRoot.recomputeViewAttributes(View;)
-> android.view.ViewRoot.checkThread()
\end{alltt}
\end{CodeOut}
%\hspace{5mm}\CodeIn{\scriptsize{org.mozilla.gecko.gfx.LayerView.<init>(Context;LayerController)}}\\
%$\rightarrow$\CodeIn{ \scriptsize{android.opengl.GLSurfaceView.setRenderer(GLSurfaceView\$Renderer;)}}\\
%$\rightarrow$\CodeIn{ \scriptsize{java.lang.Thread.start()}}\\
%$\rightarrow$\CodeIn{ \scriptsize{android.opengl.GLSurfaceView\$GLThread.run()}}\\
%$\rightarrow$\CodeIn{ \scriptsize{android.opengl.GLSurfaceView\$GLThread.guardedRun()}}\\
%$\rightarrow$\CodeIn{ \scriptsize{org.mozilla.gecko.gfx.LayerRenderer.onSurfaceChanged(GL10;II)}}\\
%$\rightarrow$\CodeIn{ \scriptsize{org.mozilla.gecko.gfx.LayerView.setViewportSize(IntSize;)}}\\
%\hspace{5mm}\CodeIn{...}\\
%$\rightarrow$\CodeIn{ \scriptsize{android.view.ViewRoot.checkThread()}}
\tinystep
\vspace*{-3.0ex} \Caption{{\label{fig:report} Our tool reports
a method call chain that reveals the potential error in
Figure~\ref{fig:androiderror}. $\rightarrow$ represents
the call relationship
between methods, and \CodeIn{checkThread} is an Android library
method  that checks whether the current thread is the
event dispatching thread before accessing a GUI object.
8 more methods in the call chain, shown as
``\CodeIn{\small{...}}" above, are omitted for brevity.
}} %\vspace{-5mm}
\end{figure}

@@

Our static analysis is independent of the call graph construction algorithm.
%Given a sound call graph, our analysis is sound in that it does
%not miss true positives.
However, modern GUI applications tend to use reflection, and
in the presence of reflection, existing call graph construction algorithms such as RTA~\cite{rta}
and k-CFA~\cite{kcfa} fail to build a complete call graph.
To alleviate this problem, we present an algorithm to build a reflection-aware call graph, and
also compare its usefulness with existing call graph construction algorithms in
our experiments.

Static analysis may report false positives due to its
conservative nature, or multiple warnings that actually correspond
to the same error. To address such limitations, we devise a set
 of error filters to remove likely false positives and redundant warnings.
The filters introduce potential unsoundness to our algorithm, but in practice
they work well and make our technique usable.

@@

We implemented an invalid-thread-error detection tool that supports
four popular GUI frameworks: SWT, Eclipse plugin environment,
Swing, and Android. Swing and SWT are the
two dominant GUI frameworks for desktop Java applications.
Eclipse is the de-facto standard IDE for Java. Android is the \#1 platform for
mobile applications with market share 56\% as of Sep 2011.
Although our technique is applicable to any GUI
framework with the single-GUI-thread rule, each framework
has its own definition of \textit{UI thread} and \textit{program 
entry points}. Thus, our implementation is parameterized with respect to
those framework-specific parts (see Section~\ref{sec:platforms}).

Our tool works in an automatic manner,
and scales to realistic programs.
We evaluated our implementation on \subnum programs comprising \totaloc LOC.
The experimental results demonstrate
that: our technique is effective (it found \bugs real-world errors
 and produced only \falses false positive warnings);
our reflection-aware call graph construction algorithm helps in
finding errors in Android applications;
and our proposed filters significantly reduce the number of warnings.
%false positives and redundant warnings.

@@

\item \textbf{Technique.} We formulate finding
the invalid thread access error as a call graph reachability problem,
and present a general error detection technique.
In addition, we use a reflection-aware
call graph construction algorithm (Section~\ref{sec:technique}).
